GWAS datasets for FDR benchmarking

### 2017-09-28 datasets/GWAS/GWAS.Rmd
Add UpSet plots for GWAS results (Issue #13)

# wrangling & plotting tools
library(data.table)
library(readxl)
library(dplyr)
library(ggplot2)
library(cowplot)

# benchmark methods
library(IHW)
library(ashr)
library(qvalue)
library(swfdr)
library(fdrtool)
```

@@

```{r, GWAS1-run, results="hide"}
cutoffs <- c(0.01, 0.025, 0.05, 0.10, 0.20)
resfile <- paste0(resdir, "bmi_results_",
                  nrow(bmi), ".RData")
if (!file.exists(resfile)){
  t1 <- proc.time()
  bmi_results <- run_benchmarks(dat=bmi, alphas=cutoffs)
  save(bmi_results, file=resfile)
  message(paste0("Took ", round((proc.time()-t1)[3],1),
                 " seconds for ", nrow(bmi), " SNPs."))
}else{
  load(resfile)
}

# plot N rejects vs FDR cutoff
p <- ggplot(bmi_results, aes(alpha, n_rejects, color=method)) +
  geom_point() +
  geom_line() +
  xlab("FDR cutoff") + ylab("Number of rejections") +
  ggtitle("BMI GWAS Metanalysis with Sample Size Covariate") +
  theme_classic()
print(p)

@@

@@

```{r, GWAS2-run, results="hide"}
# use same cutoffs as for the bmi study (defined above)
resfile <- paste0(resdir, "asthma_results_",
                  nrow(asthma), ".RData")
if (!file.exists(resfile)){
  t1 <- proc.time()
  asthma_results <- run_benchmarks(dat=asthma, alphas=cutoffs)
  save(asthma_results, file=resfile)
  message(paste0("Took ", round((proc.time()-t1)[3],1),
                 " seconds for ", nrow(asthma), " SNPs."))
}else{
  load(resfile)
}

# plot N rejects vs FDR cutoff
p <- ggplot(asthma_results, aes(alpha, n_rejects, color=method)) +
  geom_point() +
  geom_line() +
  xlab("FDR cutoff") + ylab("Number of rejections") +
  ggtitle("Asthma GWAS Metanalysis with Sample Size Covariate") +
  theme_classic()
print(p)

@@

@@

```{r, GWAS3-run, results="hide"}
# use same cutoffs as for the bmi study (defined above)
resfile <- paste0(resdir, "trig_results_",
                  nrow(trig), ".RData")
if (!file.exists(resfile)){
  t1 <- proc.time()
  trig_results <- run_benchmarks(dat=trig, alphas=cutoffs)
  save(trig_results, file=resfile)
  message(paste0("Took ", round((proc.time()-t1)[3],1),
                 " seconds for ", nrow(trig), " SNPs."))
}else{
  load(resfile)
}

# plot N rejects vs FDR cutoff
p <- ggplot(trig_results, aes(alpha, n_rejects, color=method)) +
  geom_point() +
  geom_line() +
  xlab("FDR cutoff") + ylab("Number of rejections") +
  ggtitle("Triglyceride GWAS Metanalysis with Sample Size Covariate") +
  theme_classic()
print(p)

@@

>>>>

# wrangling & plotting tools
library(data.table)
library(readxl)
library(dplyr)
library(ggplot2)
library(cowplot)
library(UpSetR)

# benchmark methods
library(IHW)
library(ashr)
library(qvalue)
library(swfdr)
library(fdrtool)

# constant parameters (for UpSet Plots)
alpha.cutoff <- 0.05
```

@@

```{r, GWAS1-run, results="hide"}
cutoffs <- c(0.01, 0.025, 0.05, 0.10, 0.20)
resfile <- paste0(resdir, "bmi_results_",
                  nrow(bmi), ".RData")
if (!file.exists(resfile)){
  t1 <- proc.time()
  bmi_results <- run_benchmarks(dat=bmi, alphas=cutoffs, pvals=TRUE)
  save(bmi_results, file=resfile)
  message(paste0("Took ", round((proc.time()-t1)[3],1),
                 " seconds for ", nrow(bmi), " SNPs."))
}else{
  load(resfile)
}

# plot N rejects vs FDR cutoff
p <- ggplot(bmi_results$stats, aes(alpha, n_rejects, color=method)) +
  geom_point() +
  geom_line() +
  xlab("FDR cutoff") + ylab("Number of rejections") +
  ggtitle("BMI GWAS Metanalysis with Sample Size Covariate") +
  theme_classic()
print(p)

@@

Next, we look beyond the number of rejections for each method and examine
the overlap between them. We'll use the Venn diagram alternative provided in the
UpsetR package.

```{r, GWAS1-upset, fig.width=9, fig.height=9}
sig_df <- bmi_results$pvals %>%
            mutate_all(function(x) 1*(x < alpha.cutoff))
upset(sig_df, nsets = 9, nintersects = 30, mb.ratio = c(0.5, 0.5),
      order.by = c("freq"), decreasing = c(TRUE), 
      sets.bar.color = "#56B4E9")
```

@@

```{r, GWAS2-run, results="hide"}
# use same cutoffs as for the bmi study (defined above)
resfile <- paste0(resdir, "asthma_results_",
                  nrow(asthma), ".RData")
if (!file.exists(resfile)){
  t1 <- proc.time()
  asthma_results <- run_benchmarks(dat=asthma, alphas=cutoffs, pvals=TRUE)
  save(asthma_results, file=resfile)
  message(paste0("Took ", round((proc.time()-t1)[3],1),
                 " seconds for ", nrow(asthma), " SNPs."))
}else{
  load(resfile)
}

# plot N rejects vs FDR cutoff
p <- ggplot(asthma_results$stats, aes(alpha, n_rejects, color=method)) +
  geom_point() +
  geom_line() +
  xlab("FDR cutoff") + ylab("Number of rejections") +
  ggtitle("Asthma GWAS Metanalysis with Sample Size Covariate") +
  theme_classic()
print(p)

@@

Next, we look beyond the number of rejections for each method and examine
the overlap between them. We'll use the Venn diagram alternative provided in the
UpsetR package.

```{r, GWAS2-upset, fig.width=9, fig.height=9}
sig_df <- asthma_results$pvals %>%
            mutate_all(function(x) 1*(x < alpha.cutoff))
upset(sig_df, nsets = 9, nintersects = 30, mb.ratio = c(0.5, 0.5),
      order.by = c("freq"), decreasing = c(TRUE), 
      sets.bar.color = "#56B4E9")
```

@@

```{r, GWAS3-run, results="hide"}
# use same cutoffs as for the bmi study (defined above)
resfile <- paste0(resdir, "trig_results_",
                  nrow(trig), ".RData")
if (!file.exists(resfile)){
  t1 <- proc.time()
  trig_results <- run_benchmarks(dat=trig, alphas=cutoffs, pvals=TRUE)
  save(trig_results, file=resfile)
  message(paste0("Took ", round((proc.time()-t1)[3],1),
                 " seconds for ", nrow(trig), " SNPs."))
}else{
  load(resfile)
}

# plot N rejects vs FDR cutoff
p <- ggplot(trig_results$stats, aes(alpha, n_rejects, color=method)) +
  geom_point() +
  geom_line() +
  xlab("FDR cutoff") + ylab("Number of rejections") +
  ggtitle("Triglyceride GWAS Metanalysis with Sample Size Covariate") +
  theme_classic()
print(p)

@@

Next, we look beyond the number of rejections for each method and examine
the overlap between them. We'll use the Venn diagram alternative provided in the
UpsetR package.

```{r, GWAS3-upset, fig.width=9, fig.height=9}
sig_df <- trig_results$pvals %>%
            mutate_all(function(x) 1*(x < alpha.cutoff))
upset(sig_df, nsets = 9, nintersects = 30, mb.ratio = c(0.5, 0.5),
      order.by = c("freq"), decreasing = c(TRUE), 
      sets.bar.color = "#56B4E9")
```
